---
title: "hw06_data_wrangling"
author: "Alejandra Urcelay"
date: "09/11/2018"
output: 
  html_document:
    toc: yes
always_allow_html: yes
editor_options: 
  chunk_output_type: console
---

## Data wrangling wrap up 

Data wrangling is the process of cleaning, structuring and enriching data into a desired format (Trifacta, 2018).

In this assignment I chose two topics to work with to improve my data wrangling skills. In the first one I completed several exercises related to strings, and in the second one I worked with a dataset to extract and analyze geographical information. 

## Topic 1: Character data

For this task I completed exercises from the [Strings chapter](https://r4ds.had.co.nz/strings.html) of R for Data Science.

I mostly worked with `stringr::words` and `stringr::sentences`

### 14.2 String basics

#### 14.2.5 Exercises

3. Use `str_length()` and `str_sub()` to extract the middle character from a string. What will you do if the string has an even number of characters?

```{r, warning=FALSE, message = FALSE}
library(tidyverse)
library(stringr)

string1 <- "abc"
string2 <- "abcd"

str_sub(string1, floor((str_length(string1)+1)/2), ceiling((str_length(string1)+1)/2))

str_sub(string2, floor((str_length(string2)+1)/2), ceiling((str_length(string2)+1)/2)) #returns the two middle characters as the string has an even number. 

```

I chose to extract both of the middle characters when having an even number, but I could also extract only one arbitrarily:

```{r}
str_sub(string2, ceiling(str_length(string2)/2), ceiling(str_length(string2)/2))
```

This way, I extract only one of the middle characters. 

6. Write a function that turns (e.g.) a vector c("a", "b", "c") into the string a, b, and c. Think carefully about what it should do if given a vector of length 0, 1, or 2.

```{r}
st_comma <- function (x, delim = ",") {
  num <- length(x)
  if(num == 0) {
   stop("vector length = 0") #error message when trying with a length 0 vector
  } else if(num == 1) {
    x
  } else if(num == 2) {
    str_c(x[[1]], "and", x[[2]], sep = " ")
  } else {
   str_1 <- str_c(x[seq_len(num - 1)], delim) #all but the last
   str_2 <- str_c("and", x[[num]], sep = " ")
   str_c(c(str_1, str_2), collapse = " ")
  }
}

#st_comma(c()) # as vector is length 0, the function throws an error message "vector length = 0"
st_comma("a") 
st_comma(c("a", "b"))
st_comma(c("a", "b", "c"))
```

### 14.3 Matching patterns with regular expressions

#### 14.3.1.1 Basic matches

2. How would you match the sequence `"'\`?

```{r}
str_view("\"'\\", "\"'\\\\")
```

3. What patterns will the regular expression `\..\..\..` match? How would you represent it as a string?

```{r}
str_view("w.x.y.z", "\\..\\..\\..")
```

It matches patterns with a dot followed by a character that repeats three times. 

#### 14.3.2.1 Anchors

1. How would you match the literal string `"$^$"` ?

```{r}
str_view("$^$", "^\\$\\^\\$")
```

2. Given the corpus of common words in `stringr::words`, create regular expressions that find all words that:

* Start with “y”.
* End with “x”
* Are exactly three letters long. (Don’t cheat by using `str_length()`!)
* Have seven letters or more.

Since the list is long, I used a match argument to show only the matching words.
When the output is too long, I use `str_subset` instead of `str_view` to have a more compact output. 

```{r}
str_view_match <- function(words, pattern) {
    str_view(words, pattern, match=TRUE)
} #function to only show matches

# start with y
str_view_match(words, "^y") 

# end with x
str_view_match(words, "x$")

# have exactly three letters
str_subset(words, "^...$")

# ≥ 7 letters
str_subset(words, ".......")
```

#### 14.3.3.1 Character classes and alternatives

1. Create regular expressions to find all words that:

* Start with a vowel.
* That only contain consonants. (Hint: thinking about matching “not”-vowels.)
* End with ed, but not with eed.
* End with ing or ise.

```{r}
# start with a vowel
str_subset(words, "^[aeiou]")

# only consonants
str_view_match(words, "^[^aeiou]+$")

# end with ed, but not with eed
str_view_match(words, "^ed$|[^e]ed$")

# end with ing or ise
str_view_match(words, "ing$|ise$")
```

2. Empirically verify the rule “i before e except after c”.

```{r}
str_view_match(words, "([^c]ie|cei)") #rule
str_view_match(words, "(cie)") #exceptions?
```

From the second output we can see there are some exceptions for this rule, such as the words `science` and `society`.


3. Is “q” always followed by a “u”?

```{r}
str_view_match(words, "q[^u]")
```

There were no words in the output, so "q" is always followed by a "u".


4. Write a regular expression that matches a word if it’s probably written in British English, not American English.

```{r}
str_view_match(words, "ise$|our") #ise instead of ize and our instead of or
```


5. Create a regular expression that will match telephone numbers as commonly written in your country

```{r}
phones <- c("(55)32498722", "(778)952-5873")
str_view(phones, "\\(\\d{2}\\)\\d{8}", match = T)
```

The output matches the telephone number as commonly written in Mexico.

#### 14.3.4.1 Repetition

3. Create regular expressions to find all words that:

* Start with three consonants.
* Have three or more vowels in a row.
* Have two or more vowel-consonant pairs in a row.

```{r}
# start with 3 consonants
str_view_match(words, "^[^aeiou]{3}")

# ≥ 3 vowels in a row
str_view_match(words, "[aeiou]{3,}")

# ≥ 2 vowel-consonant pairs in a row
str_subset(words, "([aeiou][^aeiou]){2,}")
```

#### 14.3.5.1 Grouping and backreferences

2. Construct regular expressions to match words that:

* Start and end with the same character.
* Contain a repeated pair of letters (e.g. “church” contains “ch” repeated twice.)
* Contain one letter repeated in at least three places (e.g. “eleven” contains three “e”s.)

```{r}
#start + end same character
str_view_match(words, "^(.).*\\1$")

#repeated pair of letters
str_view_match(words, "(..).*\\1")

#repeated letters
str_view_match(words, "(.).*\\1.*\\1")
```

### 14.4 Tools

#### 14.4.2 Detect matches

1. For each of the following challenges, try solving by using both a single regular expression, and a combination of multiple `str_detect()` calls.

* Find all words that start or end with x.
* Find all words that start with a vowel and end with a consonant.
* Are there any words that contain at least one of each different vowel?

```{r}
# start or end with x 
str_view_match(words, "^x|x$") #single regex
# multiple str_detect() calls
start_x <- str_detect(words, "^x")
end_x <- str_detect(words, "x$")
words[start_x | end_x]

#start with vowel end with consonant
str_subset(words, "^[aeiou].*[^aeiou]$") #single regex
#multiple str_detect()
start_vowel <- str_detect(words, "^[aeiou]")
end_cons <- str_detect(words, "[^aeiou]$")
words[start_vowel & end_cons]

#one of each vowel
allv <- c("aeioux", "aei") #to check it works
str_subset(allv, "a.*e.*i.*o.*u")
str_subset(words, "a.*e.*i.*o.*u")
# multiple str_detect()
words[str_detect(words, "a") & str_detect(words, "e") &
        str_detect(words, "i") & str_detect(words, "o") &
        str_detect(words, "u")]
```

There are no words in `stringr::words` that contain all the vowels.

2. What word has the highest number of vowels? What word has the highest proportion of vowels? 

```{r}
#highest number of vowels
num_v <- str_count(words, "[aeiou]")
max_v <- max(num_v)
words[num_v == max_v]
#highest proportion of vowels
prop_v <- str_count(words, "[aeiou]") / str_length(words)
max_p <- max(prop_v)
words[prop_v == max_p]
```

8 words have 5 vowels, which is the maximum number of values among these words. 

The word `a` has the highest proportion since length = 1 and num_v = 1.

#### 14.4.3.1 Extract matches

2. From the Harvard sentences data, extract:

* The first word from each sentence.
* All words ending in ing.
* All plurals.

```{r}
#first word
str_extract(sentences, "[^ ]+") %>% head()
sentences %>% head() #to check it worked

#words ending in ing
pat_ing <- "[A-Za-z]+ing" #define pattern
ing <- str_detect(sentences, pat_ing)
str_extract_all(sentences[ing], pat_ing) %>%
  unlist() %>%
  unique() #don't show repeated words

#plurals
str_extract_all(sentences, "[A-Za-z]{3,}s") %>%
  unlist() %>%
  unique() %>%
  head()
```

#### 14.4.4.1 Grouped matches

1. Find all words that come after a “number” like “one”, “two”, “three” etc. Pull out both the number and the word.

```{r}
pat_num <- "(one|two|three|four|five|six|seven|eight|nine|ten) ([^ ]+)"
sen_num <- sentences %>% str_subset(pat_num)
sen_num %>% str_match(pat_num)
```


2. Find all contractions. Separate out the pieces before and after the apostrophe.

```{r}
cont <- "([A-Za-z]+)'([A-Za-z]+)"
sen_cont <- sentences %>% str_subset(cont)
sen_cont %>% str_match(cont)
```

#### 14.4.5.1 Replacing matches

1. Replace all forward slashes in a string with backslashes.

```{r}
for_slash <- ("one/two/three")
str_replace_all(for_slash, "/", "\\\\") %>% writeLines()
```

2. Implement a simple version of `str_to_lower()` using `replace_all()`.

```{r}
caps <- ("ABCDE")
str_replace_all(caps, "([A-Z])", tolower)
```

#### 14.4.6.1 Splitting

1. Split up a string like "apples, pears, and bananas" into individual components.

```{r}
fruity <- ("apples, pears, and bananas")
str_split(fruity, ", and |,")
```


2. Why is it better to split up by `boundary("word")` than " "?

```{r}
fruity2 <- ("fruit: apples, pears, (bananas), and oranges")
str_split(fruity2, " ")
str_split(fruity2, boundary("word"))
```

Splitting up with `boundary("word")` is better so I don't have to specify each special punctuation character to keep out like `:` `,` or `()`.

### 14.5 Other types of pattern

1. How would you find all strings containing `\` with `regex()` vs. with `fixed()`?

```{r}
strings <- c("ab", "0\\1", "x\\y")
#regex()
str_subset(strings, regex("\\\\")) 
#fixed()
str_subset(strings, fixed("\\"))
```

2. What are the five most common words in sentences?

```{r}
(words_sen <- str_split(sentences, boundary("word")) %>%
  unlist() %>%
  str_to_lower() %>% #avoid repeated words in caps and lower
  as.tibble() %>%
  set_names("word") %>%
  group_by(word) %>%
  count(sort = TRUE) %>% #order by number
  head(5)) #only top 5
```



## Topic 4: Work with the singer data

### 4.1 Use `purrr` to map latitude and longitude into human readable information on the band’s origin places.

Notice that `revgeocode(... , output = "more")` outputs a data frame, while `revgeocode(... , output = "address")` returns a string: you have the option of dealing with nested data frames.
You will need to pay attention to two things:

* Not all of the track have a latitude and longitude: what can we do with the missing information? (filtering, …)
* Not all of the time we make a research through `revgeocode()` we get a result. What can we do to avoid those errors to bite us? (look at possibly() in purrr…)

First, I need to load the necessary packages and register my google API key to use `ggmap()`.

```{r, warning=FALSE, message = FALSE}
library(tidyverse)
library(devtools)
install_github("dkahle/ggmap")
library(ggplot2)
library(ggmap)
register_google("AIzaSyCZDZzfVa_lzlESRafTcuwH1BzethDNdcQ")
```

The data set `singer_locations` contains information about songs and associated artists in the [Million Song Dataset](https://labrosa.ee.columbia.edu/millionsong/).

Let's look at this data frame: 

```{r}
library(singer)
str(singer_locations)
library(kableExtra)
singer_locations %>% head() %>%
  kable() %>%
  kable_styling(full_width = F, position = "center")
```

The `singer_locations` data frame contains geographical information associated with the artist location stored in two different formats: 1. as a (dirty!) variable named city; 2. as a latitude / longitude pair (stored in latitude, longitude respectively).

From the output of the first songs, we can see that some tracks don't have this geographical information so I will filter to have only the ones that do contain this information.

```{r}
singer_geo <- singer_locations %>%
  filter(!is.na(city)) %>%
  select(title, artist_name, year, latitude, longitude, city) #to make table smaller
singer_geo %>%
  head() %>%
  kable() %>%
  kable_styling(full_width = F, position = "center")

nrow(singer_locations)
nrow(singer_geo)
```

After, filtering the new data frame `singer_geo` has 4129 observations, compared to 10100 that the original data frame had.
However, as there are still so many observations, I will only work with the first 25 songs.
As there are many variables too, I will only keep a few to have an easier data set to look at.

```{r}
singer_geo <- singer_geo[1:25,]
singer_geo %>%
  kable() %>%
  kable_styling(full_width = F, position = "center")
```


```{r}

singer_address <- mapply(FUN = function(longitude, latitude) { 
  revgeocode(c(longitude, latitude), output = "address")}, 
  singer_geo$longitude, singer_geo$latitude)

singer_address
```

Now `singer_address` contains the corresponding addresses from the given coordinates.
Let's see if these addresses match with the variable `city`.

### 4.2 Try to check wether the place in city corresponds to the information you retrieved.

```{r}
sing_add_city <- data.frame(address = singer_address, city = singer_geo$city)
sing_add_city %>% 
  kable() %>%
  kable_styling(full_width = F)
```

From the table we can visually compare the cities, but let's try with some code:

```{r}
library(stringi)
stri_detect_fixed(sing_add_city$address, sing_add_city$city)
```

From the output, we can see that 8 of the observations don't match. But some of these mismatches are because some cities/countries are in capitals in the column `city`.

I will try this again, putting all in low case.

```{r}
low_address <- str_to_lower(sing_add_city$address)
low_city <- str_to_lower(sing_add_city$city)
words_address <- str_split(low_address, boundary("word"))
words_city <- str_split(low_city, boundary("word"))

no_match<-function(match_length){
  match_length > 0 #to see if intersects or not
}

mapply(intersect, words_address, words_city) %>%
  lapply(length) %>%
  map(no_match)
```

After applying the previous code, the mismatches were reduced from 8 to 3. Observation number 15 is a true mismatch, since address = San Francisco, and city = New York. The other two cases are because in the column `city` only appears the state, and in the column address this state is abbreviated so they don't match. 

This things could potentially be true to different methods (i.e. having the states abbreviated in both columns) but it may compromise the accuracy of the output.

### 4.3 Go visual

```{r}
library(leaflet)

singer_geo %>%  
  leaflet()  %>%   
  addTiles() %>%  
  addCircles(lng = singer_geo$longitude,
             lat = singer_geo$latitude,
             popup = singer_geo$artist_name, 
             color = "deeppink")
```

The map can show the artist name that corresponds to the city of each pink circle.

## References

* [What is Data Wrangling?](https://www.trifacta.com/data-wrangling/)
* [Strings](https://r4ds.had.co.nz/strings.html)
* [Google API key](https://github.com/dkahle/ggmap/issues/51#issuecomment-428947722)
* [ggmap](https://cran.r-project.org/web/packages/ggmap/ggmap.pdf)
* [Applying revgecode](https://stackoverflow.com/questions/22911642/applying-revgeocode-to-a-list-of-longitude-latitude-coordinates)
* [Test if charachters are in a string](https://stackoverflow.com/questions/10128617/test-if-characters-are-in-a-string)




